\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2014-05-01 paper.tex
Rewrite of section I

\section{The challenge of human-robot interaction}

Human-robot interaction is a challenge for artificial intelligence. This field
indeed lays at the crossroad of several other domains of AI and requires to
tackle them in a holistic manner: Modeling humans and human cognition;
acquiring, representing, manipulating in a tractable way abstract knowledge;
reasoning on this knowledge to make decisions; and eventually instantiating
those decisions into physical actions in coordination with humans. Many AI
techniques are invited, from visual processing to symbolic reasoning, from task
planning to \emph{theory of mind} building, from reactive control to action
recognition and learning.

We do not tackle all these issues. This article attempts however to organize
them into a coherent challenge for artificial intelligence, and also to make
explicit some of the paths that have been investigated on our robots and that
result in a deliberative, knowledge-oriented, architecture for human-robot
interaction.

The class of interactions we consider are essentially multi-modal situated
dialog and collaborative task achievement: a robot which is expected to achieve
interactive object manipulation, fetch and carry tasks and similar tasks in a
domestic environment, with human partners (Figure~\ref{fig:hri-dec}). Two
agents (a human and a robot) share a common space and exchange information
through multiple modalities. The robot is expected to explicitly take into
account the human and to manage its interactions with him.

Namely, the robot must be able to recognize, understand and participate
to communication situations, both explicit (\eg the human addresses verbally
the robot) and implicit (\eg the human points to an object); the robot must be
able to take part to joint actions, both proactively (by planning and proposing
such an action to the human) and reactively; the robot must be able to move and
act in a safe, efficient and legible way, taking into account social rules like
proxemics.

\begin{figure}[htb]
\centering
\includegraphics[width=0.6\columnwidth]{grounding_robot.pdf}
\caption{Robot reasoning about HRI and anticipation of human activities.
  Sources of information are multi-modal dialogue, and monitoring of the
  environment and human activity. The robot must adapt on-line its behaviors
  by merging computed plans with reactive control. The robot explicitly reasons
  on the fact that it is (or not) observed by the human. Reasoning and planning
  take place at symbolic as well as geometric level and systematically take
  into account both agents beliefs and activities (from the point of view of
  the robot).}

@@

Those three items have broad implications regarding the required underlying
cognitive skills.  A \emph{joint action}, for instance, can be described as
follow. Given:

\begin {itemize}
\item a joint goal, which has been previously established and agreed
  upon (typically through dialogue),
\item the current situation, acquired through perception or
  deduction from previous perceptions, including the state of the
  environment of the robot and of the human,
\item {\it a priori} common-sense knowledge,
\end {itemize}

the robot controller computes an action to execute and who (the 
human or the robot, or both in case of a joint action) has to perform
it, and then controls or monitors its execution. The operation
continues until the goal is achieved, is declared unachievable or is
abandoned by the human.

To do so, the robot has to be equipped with a number of decisional, planning
and interpretation abilities where its human partner is taken explicitly into
account. It needs to be able:

\begin{itemize}
\item to build and maintain relevant robot and human beliefs
  (from the robot perspective) with respect to state of the world and the task,
\item to build and maintain iteratively shared (human-robot) plans, 
\item to refine and execute the actions it has to perform, and to monitor 
those achieved by its human partner.
\end{itemize}

Besides, we would like to build such abilities in a generic way, and
to provide several levels of parametrization allowing to adapt to
various environments, and various levels of involvement of the robot
ranging from teammate behavior to assistant or proactive helper.

We will see, in the following sections, how cognitive skills for realistic
human-robot interaction can be implemented and organized together in a coherent
architecture.

@@

Regarding the anchoring question, this architecture is bidirectional. The
components we described provide a \textit{bottom-up} grounding process:
geometric reasoning and dialogue processing modules constantly build and push
new symbolic contents about the world to the knowledge base where it becomes
accessible to decisional layers. In parallel, the knowledge base relies on
reasoning in a \textit{top-down} way to produce new facts that may in return
trigger physical behaviors.

@@

@@

\begin{table}
\begin{center}
\begin{tabular}{l}
\hline
Robot's beliefs about itself (\emph{robot's model}):\\
\hline
  \hspace{0.7cm}\stmt{PINK\_TRASHBIN isReachable false}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isReachable {\bf false}}\\
  \hspace{0.7cm}\stmt{LOTR\_TAPE isReachable {\bf true}}\\
  \hspace{0.7cm}\stmt{GREY\_TAPE isReachable {\bf true}}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{LOTR\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{GREY\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isOn TABLE}\\
  \hspace{0.7cm}\stmt{LOTR\_TAPE isOn TABLE}\\
  \hspace{0.7cm}\stmt{GREY\_TAPE isOn TABLE}\\
\hline
\hline
Robot's beliefs about the human (\emph{human's model}):\\
\hline
  \hspace{0.7cm}\stmt{PINK\_TRASHBIN isReachable false}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isReachable {\bf true}}\\
      \hspace{0.7cm}\stmt{LOTR\_TAPE isReachable {\bf false}}\\
      \hspace{0.7cm}\stmt{GREY\_TAPE isReachable {\bf false}}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{LOTR\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{GREY\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isOn TABLE}\\
  \hspace{0.7cm}\stmt{LOTR\_TAPE isOn TABLE}\\
  \hspace{0.7cm}\stmt{GREY\_TAPE isOn TABLE}\\ 
 \hline
\end{tabular}
\end{center}
\caption{Symbolic facts computed from the situation depicted in
Figure~\ref{fig:sparkSubfigures}. Note how reachability differs for the two
agents.}

@@

This allows us to write reactive robot controllers with a high level of
expressiveness: for instance, by subscribing to the event \stmt{human1 desires
?action, ?action type Look, ?action hasGoal myself}, we could trigger a
behavior when the human expresses (through dialogue, gestures...) that he
wants to look at the robot itself.

@@

The second one involves the {\sc Shary} execution controller and the HATP symbolic
task planner besides ORO and {\sc Dialogs}. In this scenario, the human and the
robot try to cooperatively remove objects from a table. behaviors and motions
are fully planned.

@@

We should also mention that modelling of human behaviors also appear in our
architecture at the level of task and motion planning with a set of
\emph{social rules}, including proxemics, action legibility or time
optimization.

@@

We have presented in this article a decisional framework for human-robot
interactive task achievement that is aimed to allow the robot not only to
accomplish its tasks but also to produce behaviors that support its engagement
vis-a-vis its human partner and to interpret human behaviors and intentions.
Together and in coherence with this framework, we have developed and
experimented various task planners and interaction schemes that allow the robot
to select and perform its tasks while taking into account explicitly the human
abilities as well as the constraints imposed by the presence of humans, their
needs and preferences.

@@

There is also extensive work to be done in order to refine the notion of ``good
shared plan" and ``good/acceptable robot behavior" in this context. There are
obviously large avenues for learning and adaptation in this context.

>>>>

\section{The Challenge of Human-Robot Interaction}

\subsection{The Human-Robot Interaction Context}

Human-robot interaction is a challenge for artificial intelligence. This field
lays at the crossroad of several domains of AI and requires to tackle them in a
holistic manner: Modeling humans and human cognition; acquiring, representing,
manipulating in a tractable way abstract knowledge at the human level; reasoning
on this knowledge to make decisions; and eventually instantiating those
decisions into physical actions both legible to and in coordination with humans.
Many AI techniques are invited, from visual processing to symbolic reasoning,
from task planning to \emph{theory of mind} building, from reactive control to
action recognition and learning.

We do not claim to address here the whole issue. This article attempts however to organize
it into a coherent challenge for artificial intelligence, and to explain and
illustrate some of the paths that we have investigated on our robots and that
result in a deliberative, knowledge-oriented, architecture for human-robot
interaction.

We focus on a specific class of interactions: collaborative task achievement
supported by multi-modal and situated communication. Figure~\ref{fig:hri-dec}
illustrates this context: the two agents share a common space and exchange
information through multiple modalities, and the robot is expected to achieve
interactive object manipulation, fetch and carry tasks and other similar
domestic tasks by taking into account, at every stages, the intentions, beliefs,
perspectives, skills of the human partner.  Namely, the robot must be able to
recognize, understand and participate to communication situations, both explicit
(\eg the human addresses verbally the robot) and implicit (\eg the human points
to an object); the robot must be able to take part to joint actions, both
proactively (by planning and proposing resulting plans to the human) and
reactively; the robot must be able to move and act in a safe, efficient and
legible way, taking into account social rules like proxemics.

\begin{figure}[htb]
\centering
\begin{tikzpicture}
    \node at (0,0) {\includegraphics[width=0.6\textwidth]{grounding_robot.png}};
    \node[circle, draw] at (-3.5,2.5) {A};
    \node[circle, draw] at (5.5,2) {D};
    \node[circle, draw] at (5.2,-1) {C};
    \node[circle, draw] at (-1,2) {B};
\end{tikzpicture}

\caption{The robot reasons and acts in domestic interaction scenarios.  The
    sources of information are multi-modal dialogue {\it (A)}, and
    perspective-aware monitoring of the environment and human activity {\it
    (B)}. The robot must adapt on-line its behaviours by merging computed plans
    with reactive control. The robot explicitly reasons on the fact that it is
    (or not) observed by the human. Reasoning and planning take place at
    symbolic {\it (C)} as well as geometric level and take into account agents
    beliefs, perspectives and capabilities {\it (D)} as estimated by the robot.}

@@

\emph{Communication}, \emph{Joint Actions}, \emph{Human-Aware Execution}: these
three challenges structure the whole field of human-robot interaction research,
and can be analysed in terms of the cognitive skills that they require.

A \emph{joint action}, for instance, builds from:

\begin {itemize}
    \item a joint \emph{goal}, which has been previously established and agreed
        upon (typically through dialogue),

    \item a physical environment, estimated through the robot's exteroceptive
        sensing capabilities, and complemented by inferences drawn from previous
        observations,

    \item a belief state that includes {\it a priori} common-sense knowledge and
        mental models of each of the interactors.

\end {itemize}

The robot controller then decides on an action to execute next and who has to
perform it, from the human, the robot or both in case of a joint action, and
eventually controls or monitors its execution. The operation continues until the
goal is achieved, is declared unachievable or is abandoned by the human.

\begin{inparaenum}
This translates into several decisional, planning, representation skills that
need to be available to the robot. It must be able \item to represent and
manipulate symbolic belief states, \item to acquire and maintain them up-to-date
with respect to state of the world and the task at hand, \item to build and
maintain, and iteratively refine shared (human-robot) plans, \item to
instantiate and execute the actions it has to perform, and conversely,
to monitor those achieved by its human partner.
\end{inparaenum}

Such abilities must be designed and implemented in generic ways and
provide several levels of parametrization such as they adapt to
various environments, different tasks and variable levels of engagement of the
robot, ranging from teammate behaviour to assistant or proactive helper.

\subsection{The Key Issues}

TDB

@@

Regarding the anchoring question, this architecture is bidirectional. The
components we described provide a \textit{bottom-up} grounding process:
geometric reasoning and dialogue processing modules constantly build and push
new symbolic contents about the world to the knowledge base where it becomes
accessible to decisional layers. In parallel, the knowledge base relies on
reasoning in a \textit{top-down} way to produce new facts that may in return
trigger physical behaviours.

@@

SPARK is designed to run and continously update the knowledge base (in our
experiments, it runs at about 10Hz). At each steps, it re-computes spatial
relationships and affordances for the whole scene, and send the delta (new
relations and relations that do not hold anymore) to the knowledge base. This
approach may raise scalability concerns (we did not however hit such issues in
our constrainted scenarios), but greatly simplifies the management of the
\emph{dynamics} of the knowledge (\emph{When do I discard outdated knowledge?
When do I update it?}). Since it is equivalent to a reset of the reasoner
domain, it also hides issues linked to non-monotonic reasoning
(see~\cite{McCarthy2007} for a discussion on that question).

@@

\begin{table}
\begin{center}
    \subfigure{
\begin{tabular}{l}
\hline
Robot's beliefs about itself (\emph{robot's model})\\
\hline
  \hspace{0.7cm}\stmt{PINK\_TRASHBIN isReachable false}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isReachable {\bf false}}\\
  \hspace{0.7cm}\stmt{LOTR\_TAPE isReachable {\bf true}}\\
  \hspace{0.7cm}\stmt{GREY\_TAPE isReachable {\bf true}}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{LOTR\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{GREY\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isOn TABLE}\\
  \hspace{0.7cm}\stmt{LOTR\_TAPE isOn TABLE}\\
  \hspace{0.7cm}\stmt{GREY\_TAPE isOn TABLE}\\
\hline
\end{tabular}
}\hspace{1em}%
\subfigure{
\begin{tabular}{l}
\hline
Robot's beliefs about the human (\emph{human's model})\\
\hline
  \hspace{0.7cm}\stmt{PINK\_TRASHBIN isReachable false}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isReachable {\bf true}}\\
      \hspace{0.7cm}\stmt{LOTR\_TAPE isReachable {\bf false}}\\
      \hspace{0.7cm}\stmt{GREY\_TAPE isReachable {\bf false}}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{LOTR\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{GREY\_TAPE isVisible true}\\
  \hspace{0.7cm}\stmt{WALLE\_TAPE isOn TABLE}\\
  \hspace{0.7cm}\stmt{LOTR\_TAPE isOn TABLE}\\
  \hspace{0.7cm}\stmt{GREY\_TAPE isOn TABLE}\\ 
 \hline
\end{tabular}
}
\end{center}
\caption{Symbolic facts computed from the situation depicted in
Figure~\ref{fig:sparkSubfigures}. Note how reachability differs for the two
agents.}

@@

This allows us to write reactive robot controllers with a high level of
expressiveness: for instance, by subscribing to the event \stmt{human1 desires
?action, ?action type Look, ?action hasGoal myself}, we could trigger a
behaviour when the human expresses (through dialogue, gestures...) that he
wants to look at the robot itself.

@@

The second one involves the {\sc Shary} execution controller and the HATP symbolic
task planner besides ORO and {\sc Dialogs}. In this scenario, the human and the
robot try to cooperatively remove objects from a table. behaviours and motions
are fully planned.

@@

We should also mention that modelling of human behaviours also appear in our
architecture at the level of task and motion planning with a set of
\emph{social rules}, including proxemics, action legibility or time
optimization.

@@

We have presented in this article a decisional framework for human-robot
interactive task achievement that is aimed to allow the robot not only to
accomplish its tasks but also to produce behaviours that support its engagement
vis-a-vis its human partner and to interpret human behaviours and intentions.
Together and in coherence with this framework, we have developed and
experimented various task planners and interaction schemes that allow the robot
to select and perform its tasks while taking into account explicitly the human
abilities as well as the constraints imposed by the presence of humans, their
needs and preferences.

@@

There is also extensive work to be done in order to refine the notion of ``good
shared plan" and ``good/acceptable robot behaviour" in this context. There are
obviously large avenues for learning and adaptation in this context.
